---
title: "Practical Machine Learning Project"
author: "JC Harrop"
date: '2017-07-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Simple version of the model process for the assignment
```{r}
library(tidyverse)
library(caret)
# library(parallel)
# library(doParallel)

# cluster <- makeCluster(detectCores()-1)
# registerDoParallel(cluster)
```

Lets just load the training data and have at it.  

```{r}
setwd("~/Documents/R work/DSS8-Project")
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
set.seed(749)  # not sure if I really need this.  Habit.
```

Inspection of the data indicated there was some room for cleaning and selecting a subset 
of parameters.  We build a working set.

```{r cleaning}
working <- training[,-seq(1:7)]
working[working==""] <- NA
indexNA <- as.vector(sapply(working[,1:ncol(working)],
                            function(x) {length(which(is.na(x))) < 1000}))
working <- working[,indexNA]

working <- working[working$gyros_dumbbell_x > (-50), ]
working <- working[working$magnet_belt_z < (-125), ]
working <- working[working$magnet_belt_x < (250), ]
working <- working[working$yaw_belt < (100), ]
working <- working[working$yaw_belt > (-100), ]
working <- working[working$magnet_dumbbell_y > (-1000), ]

working <- subset(working, 
                   select=c(-magnet_forearm_z, -gyros_forearm_z, -magnet_arm_x, 
                            -gyros_arm_y, -gyros_arm_z, -accel_belt_x, -gyros_belt_y, 
                            -total_accel_belt, -roll_belt, -roll_dumbbell,
                            -gyros_dumbbell_z, -gyros_dumbbell_x, -accel_belt_y, 
                            -magnet_belt_x, -total_accel_arm, -magnet_arm_y, 
                            -gyros_forearm_x, -gyros_forearm_y, -accel_forearm_x))

summary(training$classe)
```

Although we tried various methods to split correct and flawed exercises, we will run
all categories through the random forest algorithm rather than the two stage seperation.
